---
title: "Rprogramming notes"
author: "me"
date: "6/28/2021"
output: pdf_document
---
# Objects
create any vector by `vector()`
list can include 
explicity wants a integer suffix L 1(num) 1L(int)
`attributes()` can modify names, dimensions, class, length etc.
`c()` is a func to be used to create vectors of objects
finally , by using `vector()` "datatype", length=..
you can set the vector as you wanted
then uses `class(x)` to know the datatype currently 

```{r}
x<- c(0.5, 0.6)     ## numeric 
x<- c(TRUE, FALSE)  ## logical
x<- c(T, F)         ## logical
x<- c("a","b","c")  ## charac
x<- 9:29            ## integer
x<- c(1+0i, 2+3i)   ## complex

x <- vector("numeric", length =10)
x
class(x)
```

# mixed objects
```{r}
y <- c(2,"y") ## character
y
y <- c(TRUE, 3) ## numeric
y
y <- c("a", TRUE) ## character
y
```

# using as.datatype(x)

```{r}
x <- 0:8
x
class(x)
as.logical(x)
as.numeric(x)
as.character(x)
as.complex(x)
```


```{r}
b <- c("1","2","3")
class(b)
as.logical(b)
as.numeric(b)
```

# lists
every elements in the lists can be different classes and length
(dataframe, same length and within cols and rows)
```{r}
x <- list( TRUE, 2+3i, "g", 3)
x
```

# matrices <- vectors that have dimensions
matrix same class with cols and rows
create a matrix by 3 ways
1. `matrix(3:10, ncol=3, nrow = 4)`
2. made a sequence first, then attribute dimension
```{r}
m <- 1:15
dim(m)<-c(5, 3)
m
```
3. combine vectors into matrix
```{r}
x <- 1:4
y<- 10:13
cbind(x,y)
rbind(x,y)
```
you can use `dim()` for knowing the col and row for the matrix
and you've learned `matrix(3:10, ncol=3, nrow = 4)``cbind(a,b)``rbind(x,y)`.

# factors
factors are used for represent categorical data
```{r}
x<- factor(c("yes", "yes", "no", "yes", "no"))
x
table(x)
unclass(x)
```
`table()`can show you how many levels are there
and `unclass()` can tell you how it's calculate as a integer in R, it attribute 1 as no and 2 as yes, this follows the order of the alphabet, n go forward than y, so if you want to redefine the baseline of the factor, you may use `levels = c("yes", "no")` in the factor func.
```{r}
x <- factor(c("yes", "no", "yes", "yes", "no"), levels =c("yes","no"))
x
```

# missing values
nan ⊂ na
nan only finds nan
na finds both nan and na
```{r}
x <- c(1,2,NaN, NA,4)
is.nan(x)
is.na(x)
```

# Data frame
dataframe doesn't need same class or same length.
and everyrow have a name

```{r}
x <- data.frame(foo = 1:4, bar =c(T,T,F,F))
x
nrow(x)
ncol(x)
```
---------------error----------------- 

```{r}
a<- c("   ", "list","matrix","dataframe","same class","X","O","X", "cols and rows","X","O","O","same length","X","O","X")
dim(a)<- c(4,4)
a
```
```{r}
b <- data.frame(sameclass = ("X","O","X"),cols&rows <- c("X","O","O"), same_length <-c("X","O","X"))
b
```
---------------fixed-----------------
prac on my own
by constructing vectors forwardly, then combine them using data.frame(col = vec)
```{r}
sameclass <- c("X","O","X")
sameclass
colrow <- c("X","O","O")
colrow
same_length <-c("X","O","X")
same_length
data1<- c("list","matrix", "data frame")
data1
c <- data.frame(datatype = data1, sameclass = sameclass, colrow = colrow, same_length = same_length)
c
```

# name attribute
```{r}
x <- 1:3
names(x)
names(x) <- c("1st","2nd","3rd")
names(x)
x
```

```{r}
x <- list(x=1,y=2,c=3)
x
```

```{r}
m <- matrix(1:6, nrow = 3, ncol = 2)
m
dimnames(m) <- list(c("a","b","c"),c("one","two"))
m
```

# reading data
`read.table()`
`read.table()`
set the `comment.char = ""` if there are no commented line in my file.
`colClasses= "numeric"`will be wriiten in the read function, and it's really important, it can run FASTER
, it'll dont't need to figure out the colClasses on its own.

# dput/dget and dump()
made up a dataframe,then dput it and d get it.
unlike wrinting out a table or csv file, 
dput and dump preserve the metadata(like the pic with json, formateed), so that another user doesn't have to specify it all over again.
adv : it's textual + metadata, it won't need reconstruct

```{r}
y <- data.frame(a=1, b="a")
y
dput(y)
dput(y,file = "dputtest.R")
new.calldput <- dget("dputtest.R")
new.calldput

```
dput can only be used in 1 r object,
dump() can be used on multiple r obj.
```{r}
x <- "w"
y <- data.frame(a=1 , b = "a")
dump(c("x","y"),file = "dumptry.R")
rm(x,y)
source("dumptry.R")
y
x
```
i still cannot figure out what does source means

# reading files from external
there are `file()` `url()` `bzfile()`  `gzfile()`
don't forget close the connection after using it by `close()`.

```{r}
con <- file("midterm.csv")
x <- readLines(con, 10)
x

hey <- url("https://www.jhsph.edu","r")
x <- readLines(hey,10)
head(x)
```

# subsetting 
`[]` returns 
`[[]]`
`$`
```{r}
x <- c("a","b","c","c","d","a")
x
x[1]
x[2]
x[2:5]
x[3:7]
x[x> "b"]
u <- x > "b"
u
x[u]  
```

subsetting lists
```{r}
x <- list(foo = 1:4, bar =0.6)
x ## it'll print 2 list with name and sequence
x[1] ## the first list

## same returns : the sequence of the first list
x[[1]] 
x[["foo"]]
x$f
x[["f", exact = FALSE]]

x["bar"] ## the 2nd list
x$bar ## $ is similar to [[]]
x[[2]]
x$b
```

```{r}
x<- list(foo = 1:4, bar = 0.6, baz = "hello")
x ## can see 3 lists
x[c(2,3)] ## choosing multiple  ele
 
name <- "foo"
x[name] ## whole list
x[[name]] ## sequence only
x$name ## NULL　$ can only be used with literal name
x$foo ## sequence
```

```{r}
x <- list(a= list(11,13,15),b = c(3.14,2.81))
x ## a as a list will have index. bdoes not
x[[c(1,2)]]
x[[1]][[3]]
x[[c(2,1)]]
x[[2]][[2]]
```
subsetting matrix
```{r}
x <- matrix(1:6,2,3) ## data row col
x
x[1,2] ## row=1 col=2 but only number returns
x[1,2,drop = FALSE] ## one by one matrix
x[2, ] ## row =2 col= 
x[2, ,drop = FALSE]



```

# partial matching
a tool that saves you from typing
```{r}
x <- list(allthatbase = 1:5)
x$a
x[["a"]] ## [[]] needs the exact name, no partial mathching
x[["a", exact = FALSE]] ## Turn off the default
```

# reremoving NA values
using`is.na()` to return the logical response
and`x[!bad]` to exclude the ele that reach `is.na()`
```{r}
x <-  c(1, 2, NA , 4, NA, 5)
x
x[3]
bad <- is.na(x) ## is.na() is a function of NA logical return
bad
x[!bad]
```


`complete.cases()` is a func to record the index of complete cases in the list, watch out for using wrong index!!
```{r}
x <- c(1, 2, NA , 4, NA, 5, NA)
y <- c("a", "b", NA, "d", NA, "f", "g")

xgood <- complete.cases(x)
xgood
ygood <- complete.cases(y)
ygood
allgood <- complete.cases(x,y)
allgood

x[ygood]
x[xgood]
y[ygood]

cc <- data.frame(x = x, y = y)
cc
cc <- complete.cases(cc)
cc
```

```{r}
airquality[1,2]
airquality[1:10, ]
good <- complete.cases(airquality)
airquality[good, ][1:10, ]

data.frame(airquality = airquality[good, ])
data.frame(airquality[good, ])
bb <- data.frame(airquality[good, ])
bb[1:10,] ## without this method, how to add the length into the dataframe??
```

# quiz

16. how nany missing values are in the Ozone col of this data frame?
```{r}
x <- read.csv("hw1_data.csv")
y = subset(x, is.na(Ozone))
y
nrow(y)
```

17.What is the mean of the Ozone column in this dataset? Exclude missing values (coded as NA) from this calculation.
```{r}
z = subset(x, !is.na(Ozone), select = Ozone)
z
colMeans(z)
apply(z,2,mean) ## what does fun = 2 means?
                ##  for a matrix 1 indicates rows, 2 indicates columns
```

18.Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. What is the mean of Solar.R in this subset?
```{r}
b <- subset(x, Ozone > 30 & Temp > 90, select = Solar.R)
colMeans(b)
```

19.What is the mean of "Temp" when "Month" is equal to 6?
```{r}
c <- subset(x, Month == 6, select = Temp:Month)
colMeans(c)
```

What was the maximum ozone value in the month of May (i.e. Month = 5)?
```{r}
d <- subset(x, Month == 5 & !is.na(Ozone), select = Ozone)
## must except the NAs , max will uses the NAs
apply(d, 2, max)
```



